---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r echo = TRUE}

MyData <- read.csv('C:/Users/scott/Documents/GitHub/RepResP1/repdata-data-activity/activity.csv', header = TRUE)
MyData <- MyData[!is.na(MyData$steps), ]
MyData$date <- as.Date(MyData$date, "%Y-%m-%d")

```


## What is mean total number of steps taken per day?
```{r echo = TRUE}

meanData <- aggregate(MyData$steps, list(date=MyData$date), mean)
medianData <- aggregate(MyData$steps, list(date=MyData$date), median)

```

## What is the average daily activity pattern?
```{r echo = TRUE}

meanTS <- aggregate(MyData$steps, list(interval=MyData$interval), mean)
plot(MyData$interval, MyData$steps, type="l", xlab= "Interval", ylab= "Steps")
meanTS[which.max( meanTS[,2] ),]

```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps? 
Interval 835.

## Imputing missing values
```{r echo = TRUE}

MyNA <- read.csv('C:/Users/scott/Documents/GitHub/RepResP1/repdata-data-activity/activity.csv', header = TRUE)
MyNA$date <- as.Date(MyNA$date, "%Y-%m-%d")
MyNA <- MyNA[is.na(MyNA$steps), ]
nrow(MyNA)

MergeData <- merge(MyNA, meanTS, by="interval")[, c('x', 'date', 'interval')]
names(MergeData) <- c("steps", "date", "interval")
FullData <- rbind(MyData, MergeData)

MySum <- aggregate(FullData$steps, list(date=FullData$date), sum)
barplot(MySum$x, names.arg=MySum$date, ylab = "Steps", xlab = "Date")

meanFull <- aggregate(FullData$steps, list(date=FullData$date), mean)
medianFull <- aggregate(FullData$steps, list(date=FullData$date), median)
head(meanData[order(meanData$date),])
head(meanFull[order(meanFull$date),])
head(medianData[order(medianData$date),])
head(medianFull[order(medianFull$date),])

```

What is the impact of imputing missing data on the estimates of the total daily number of steps?  
I'm not seeing any change in the periods in both data sets.

## Are there differences in activity patterns between weekdays and weekends?
```{r echo = TRUE}

FullData$dayOfWeek <- weekdays(as.Date(FullData$date, "%Y-%m-%d"))

WeekDay <- subset(FullData, !(dayOfWeek == "Saturday" | dayOfWeek == "Sunday"))
WeekEnd <- subset(FullData, dayOfWeek == "Saturday" | dayOfWeek == "Sunday")

par(mfrow=c(2,1))
plot(WeekEnd$interval, WeekEnd$steps, type="l", xlab= "Interval", ylab= "Steps", main="WeekEnd")
plot(WeekDay$interval, WeekDay$steps, type="l", xlab= "Interval", ylab= "Steps", main="WeekDay")
grid()

```